---
title: "Introduction to lifeR"
author: "Jeffrey C. Oliver"
date: "24 August, 2021"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to lifeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lifeR)
```

# Summary

What does this package do?

# Installation

Currently, you will need to install the package from GitHub; you can do this 
with help from the remotes package:

```{r install, eval = FALSE}
install.packages("remotes")
remotes::install_github(repo = "jcoliver/lifeR")
```

If you want to have the introductory vignette included in the installation, 
then pass `build_vignettes = TRUE` in the call to `install_github()`:

```{r install-vignette, eval = FALSE}
remotes::install_github(repo = "jcoliver/lifeR", build_vignettes = TRUE)
```

# Using lifeR

## Requirements

You will need a minimum of three pieces of information to use lifeR:

1. An eBird API key
2. A list of species you have seen (life list, year list, etc.)
3. A pair of latitude and longitude coordinates to center the search for 
potential locations.

### 1. eBird API key

Because lifeR relies on the public eBird API, you will need to have your own 
unique API key to use this package. To get your key, visit [https://ebird.org/api/keygen](https://ebird.org/api/keygen) to get a eBird 
API key (note you will be prompted to log into your eBird account). The key 
will be a random string of letters and numbers. I suggest you copy the key and 
save it to a plain text file on your computer; you can use a program like 
Notepad or TextEdit to create plain text files. Save this file somewhere you 
will remember (like "My Documents") and name it something that is recognizable 
(e.g. "ebird-api-key.txt"). We will use this file later to access the eBird 
API.

### 2. Your species list

This should be a list of the species you have already seen, but it can vary 
based on the list you would like to add species to. If you want to find sites 
where birds you have never seen have been observed recently, you would use your 
life list. If you are working on a Big Year, you would use your year list. 
Trying to break 200 species for your Santa Cruz, Arizona county list? Then yes, 
use your Santa Cruz, Arizona species list. These can all be downloaded from 
your eBird account by going to 
[https://ebird.org/lifelist/](https://ebird.org/lifelist/), using the dropdown 
menus near the top of the page to select the region and time range of your 
list, then downloading the list through the "Download (csv)" link near the 
upper-right corner of the page. Save the file somewhere where you will remember 
where it is.

### 3. Latitude and longitude

Finally, lifeR is designed to identify sites in a region for you to visit. 
Therefore, you will need to provide information on _where_ that region is. The 
package uses a latitude and longitude coordinate to serve as the "center" of 
the region of interest, and find sites within a certain radius of that center. 
The maximum radius is 50 kilometers - if you want to search a larger area, you 
will just need to define two centers (information on doing this is provided 
below). If you do not know the latitude and longitude of your "center" I 
recommend Google Maps, which allows you to right- or control-click on a 
location on a map, and the latitude and longitude coordinates will be 
displayed. Note lifeR requires decimal degree formatted data (e.g. -110.92, 
_not_ 110° 55' 12").

## Create the report

### Helpful tips

+ Won't see everything
+ The longer your list, the fewer sites

## How does it work?

## What doesn't the package do?

+ Connect to your eBird account
+ Include additional measures of confidence, due to data returned by eBird